---
title: "Animal rescues in London"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(plotly)
library(leaflet)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
animal_rescues <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-29/animal_rescues.csv')

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
skimr::skim(animal_rescues)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Set the date type, latitude and longitude.
newdata = animal_rescues %>% 
  dplyr::transmute(
    hourly_notional_cost,
    year = as.factor(cal_year),
    date_time_of_call = lubridate::dmy_hm(date_time_of_call),
    animal_type = animal_group_parent %>% as.factor(),
    ward = ward %>% as.factor(),
    latitude = as.numeric(latitude), 
    longitude = as.numeric(longitude)) %>% 
  dplyr::filter(complete.cases(latitude), complete.cases(longitude), latitude > 0, longitude > 0)

```

Overview
======================================================================

Column {data-width=350}
-----------------------------------------------------------------------

### Data

```{r}
skimr::skim(animal_rescues)
```

### Most reported animal cases

<!--Which animals were reported most?-->

```{r echo=FALSE, warning=FALSE, message=FALSE}

# Get the total reported by animal type.
animals_reported <- newdata %>%
  filter(!is.na(animal_type)) %>%
  count(animal_type, sort = TRUE) 

top_animals_reported <- animals_reported %>%
  slice_head(n = 6) %>%
  arrange(n) %>%
  mutate(animal_type = factor(animal_type, levels = animal_type)) 
  
plot_ly(data = top_animals_reported,
        x = ~n,
        y = ~animal_type,
        color = ~animal_type,
        orientation = "h",
        type = "bar") %>%
  layout(title = "Reported animal cases",
         yaxis = list(title = ""),
         xaxis = list(title = "Number of cases"))

```

Column {data-width=650}
-----------------------------------------------------------------------

### Most common regions 

<!--What regions are most common?-->

```{r echo=FALSE, warning=FALSE, message=FALSE}

# Get the total reported cases by region.
regions_reported <- newdata %>%
  filter(!is.na(ward)) %>%
  count(ward, sort = TRUE) 

# Get the top 8 records.
top_regions_reported <- regions_reported %>%
  slice_head(n = 8) %>%
  arrange(n) %>%
  mutate(ward = factor(ward, levels = ward)) 

# Flipped bar plot for the top reported cases by region.
plot_ly(data = top_regions_reported,
        x = ~n,
        y = ~ward,
        color = ~ward,
        orientation = "h",
        type = "bar") %>%
  layout(title = "Reported cases by region",
         yaxis = list(title = ""),
         xaxis = list(title = "Number of cases"))

```

### Reported cases and region

```{r, echo=FALSE}

# Get the records of the top reported cases by region. 
ward_map_data <- newdata %>% 
  filter(ward %in% top_regions_reported$ward) %>%
  select(latitude, longitude, ward, animal_type) %>%
  mutate( mytext=paste(
    "City: ", ward, "\n", 
    "Animal: ", animal_type, sep=""))

```


```{r, echo=FALSE}
# Create the text.
mytext <- paste(
   "Animal: ", ward_map_data$animal_type, "<br/>", 
   "Ward: ", ward_map_data$ward, sep="")  %>%
  lapply(htmltools::HTML)
                

# Final Map
m <- leaflet(ward_map_data) %>% 
  addTiles()  %>% 
  setView( lat=51.1, lng=0 , zoom=8) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addCircleMarkers(~longitude, ~latitude, radius =1,
                   label = mytext,
                   labelOptions = labelOptions( style = list("font-weight" = "normal",
                                                             padding = "3px 8px"),
                                                             textsize = "13px", 
                                                             direction = "auto"))
   
m 
```

Analysis of cost
==========================================================================

### Notional cost by animal type

```{r, echo = FALSE}
# Notional cost by animal type.
p2 <- newdata %>%
  filter(!is.na(animal_type), !is.na(hourly_notional_cost),
         animal_type %in% c("Fox", "Deer", "Horse", "Bird", "Dog", "Cat")) %>%
  ggplot() +
  geom_boxplot(aes(x = animal_type, hourly_notional_cost, color = animal_type)) +
  geom_jitter(aes(x = animal_type, hourly_notional_cost, color = animal_type)) +
  labs(x = "Animal Type", y = "Hourly notional cost") 
 
ggplotly(p2)
```


Reported cases by year
==========================================================================

```{r, echo = FALSE}
p3 <- newdata %>%
  group_by(animal_type) %>%
  filter(animal_type %in% top_animals_reported$animal_type) %>%
  ggplot() +
  geom_bar(aes(x = year, fill = animal_type)) +
  facet_wrap(~animal_type)

ggplotly(p3)

```



About
==========================================================================

This analysis is about the reported cases of animals that were trapped and rescued by the London Fire Brigade.

The data is taken from the TidyTuesday 2021 week 27 dataset. 

You can access the original data from [here.](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-06-29/readme.md) 

Variables :

* Animal Type
* Region
* Year
* Longitude
* Latitude
* Hourly notional cost

Author: 

H. Lakna Perera

University of Jayewardenapura









